Low Complexity Adaptive Sonar Imaging

### 2018-02-19 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2016:07:28	12:02:03	(to.28.juli)
2_MVDR ok. 3_LCA inlcuded

% \RequirePackage[latin1]{inputenc}%              % Set input encoding (optionally latin1)
% \RequirePackage[T1]{fontenc}%                 % Set font encoding
% \usepackage[norsk]{babel}

% Single column review mode:
%\documentclass[12pt,journal,onecolumn]{IEEEtran}
%\newcommand\CLASSINPUTbaselinestretch{2}
%\RequirePackage{calc}
%\RequirePackage{fp} 
% Left margin = 1 inch + hoffset + oddsidemargin (or evensidemargin)
% Adding the 2cm gutter width to the odd/even side margins:
%\setlength\hoffset{0pt}
%\setlength\oddsidemargin{4cm}       % 4cm margins on the left side
%\addtolength\oddsidemargin{-1in}    % Subtract the initial 1 inch
%\setlength\textwidth{21cm-8cm}    % A4 width (21cm) minus margins on either side

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                  ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% IEEE ''APPROVED'' PACKAGES %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                  ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

% 
% \ifCLASSOPTIONcaptionsoff                       % IEEE promoted hack to turn off captions from the 
%    \let\MYorigsubfloat\subfloat                 % subfloat package should the captionsoff option
%    \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}} % be specified.
% \fi

@@

% \ifPeerReview
% \let\OldIncludegraphics{\includegraphics}
% \usepackage{letltxmacro}
% \LetLtxMacro{\OldIncludegrsaphics}{\includegraphics}
% \renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=\linewidth, #1]{#2}}
% \fi

% % \makeatletter
% % \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
% % \else\Gin@nat@width\fi}
% % \makeatother
% % \let\Oldincludegraphics\includegraphics
% % \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

@@

\title{Low Complexity Adaptive Sonar Imaging}

\author{Jo~Inge~Buskenes, %
        Roy~Edgar~Hansen, %
        Andreas~Austeng%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem All authors are with the Department of Informatics, University of Oslo, Norway.}%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem R.~E.~Hansen is also with the Norwegian Defense Research Establishment (FFI), Norway.}% <-this % stops a space

% \thanks{Manuscript received April 19, 2005; revised January 11, 2007.}
}

% The paper headers
\markboth{IEEE Journal of Oceanic Engineering}%
{Low Complexity Adaptive Beamformer for Active Sonar Imaging}

% Publishers ID mark:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
% 
% \begin{itemize}
% \item Dynamic weights $\Rightarrow$ greater contrast/resolution potential
% \item But:
% \begin{itemize}
% \item Active system $\Rightarrow$ noise and signal correlated $\Rightarrow$ adaptive beamformers no longer robust.
% \item Subarray averaging decorrelate the noise, but in the process we sacrifice resolution.
% \item Computationally complex O($M^3$) due to inversion of a spatial covariance matrix. LCA is only O($MW$). Even if Capon subspace/beamspace versions exist, they remain more complex than the LCA, and is not as ideally suited for implementation of SIMD hardware (GPUs, DSPs, ...).
% \end{itemize}
% \item LCA is inherently robust, is ``parameter free'' once setup with a well designed window set, and can often be implemented with minor modifications to already existing DAS beamformers.
% 
% The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels.
% 
% While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output. 
% 
We have studied the low complexity adaptive (LCA) beamformer in active sonar imaging. LCA can be viewed as either a simplification of the minimum variance distortionless response (MVDR) beamformer, or as an adaptive extension to the delay and sum (DAS) beamformer. While both LCA and MVDR attempt to minimize the power of noise and interference in the image, MVDR achieves this by computing optimal array weights from the spatial statistics of the wavefield, while LCA selects the best performing weights out of a predefined set.

To build confidence in the LCA method we show that a robust MVDR implementation typically creates weights spanning between the rectangular and Hamming window function. We let LCA select from a set of Kaiser windows with responses in this span, and add some steered variations of each. We limit the steering to roughly half the \minus{}3\,dB width of the window's amplitude response. Using experimental data from the Kongsberg Maritime HISAS1030 sonar we find that LCA and MVDR produce nearly identical images of large scenes, both being superior to DAS. On point targets LCA is able to double the resolution compared to DAS, or provide half that of MVDR. This performance is achieved with a total of 6 windows; the rectangular window and the Kaiser window with $\beta=5$, in an unsteered version, and versions that are left and right steered to the steering limit. Slightly smoother images are produced if the window count is increased to 15, but past this we observe minimal difference. Finally we show that LCA works just as well if Kaiser windows are substituted with trigonometric ones.

All our observations and experiences point to LCA being very easy to understand and manage. It simply works, and is surprisingly insensitive to the exact type of window function, steering amount, or number of windows. It can be efficiently implemented on parallel hardware, and handles any scene without the need for parameter adjustments.

% An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.

% \end{itemize}

\end{abstract}

% Keywords (normally not used for peer reviews)
\ifPeerReview\else
\begin{IEEEkeywords}
Beamforming, adaptive beamforming, MVDR, LCA, sonar, active, complexity.
\end{IEEEkeywords}
\fi
}
% \fi

% make the title area
\maketitle

% This command fixes abstract positioning for compsoc articles:
\IEEEdisplaynotcompsoctitleabstractindextext

% (Optional) Add some extra info on cover page of peer review papers:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi

% Insert page break and insert second title (peer review mode)
\IEEEpeerreviewmaketitle

\section{Introduction}

% - Nature and scope
% - Past work
% - Methods
% - Brief summary results
% - Brief conclusions

% The promise is already demonstrated in nature.
% 
% The sonar systems of the future 
% 
% Future sonar systems 
% 
% Great promise in adaptivity
% 
% The greatest untapped 
% 
% The sonars of the future might benefit greatly from adaptivity.
% 
% 
% Future sonars will likely be

% To make a sonar truly perform better, we need to make it smarter. 
% 
% \IEEEPARstart{F}{uture} sonar systems will merit a much more sophisticated level of adaptivity. 
% 
% \IEEEPARstart{T}{he} next generation sonar systems will be adaptive.
% 
% Making sonars adaptive is the way to take them into the future. 
% 
% The virtue of future sonar systems will be sophisticated adaptivity. 
% 
% Intelligent
% 
% Reactive
% 
% Observe, process, adapt, act. 
% 
% Adaptivity - sense, process, respond
% 
% Faster, smarter
% 
% The best sonars in nature are adaptive. 
% Nature allures us to be this bold because

\IEEEPARstart{T}{he} best sonars in existence is perhaps found in nature. Bats, for instance, use a high frequency sonar to detect, identify and track prey with amazing precision in caves amongst a myriad of other bats. No man made sonar can currenly match this feat, much due to the challenge of processing and adapting to such a complex environment fast enough. This is why human made sonars have little to no adaptivity on transmit. However, several methods have been investigated to adaptively post process the received data. Perhaps the best studied method for reconstructing a sonar image out of the data is the minimum variance distortion-less response (MVDR) beamformer~\cite{Capon1969}. It adapts the sonar array's spatial response to minimize the influence of noise and interference in the final image.

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This is because MVDR relies on estimating and inverting a covariance matrix. The estimation step dominates the computation in sonar systems with less than 32 channels, but here the computational complexity can be reduced and the implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014}. For larger systems the inversion step dominates. This can be dealt with by approximating the full covariance matrix with a small one using space reduction techniques such as beamspace processing~\cite{Asen2013}, or the closely related principal component analysis (PCA) method~\cite{Kim2014}. Another alternative is to assume spatial stationarity to form more easily invertible Toepliz matrices~\cite{Asl2012}. However, the mentioned methods are still relatively slow compared to DAS.

A much faster and simpler alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. Based on an idea by Vignon~\cite{Vignon2008}, it was first introduced by Synnev\aa{}g \emph{et al}. in clinical medical ultrasound imaging~\cite{Synnevag2011}, who demonstrated its ability to obtain very similar image quality to MVDR in systems with focused transmit beams. LCA applies a set of predefined and static windows, or apodization weights, and selects the one offering the best noise suppression. It relies on the same optimization criterion as MVDR, thus may be considered a version of MVDR with a discrete and static window solution space. Alternatively, it may be viewed as a multi-apodization technique. One such method is described by Stankwitz \emph{et al}. in radar imaging~\cite{Stankwitz1995}, where the best out of 2 or 3 windows are selected. However, LCA is different in that it typically selects from a larger pool of windows and allows phase-steered variations of them. This gives LCA an adaptive freedom that lies somewhere in between that of traditional multi-apodization techniques and MVDR. How similarly LCA performs compared to MVDR depends on how well the predefined windows represent the constrained solution space of MVDR.

% While using weights that span a bigger space gives LCA greater freedom, not The bigger the space the weights span, the greater the adaptive freedom, but only the solutions that are valid in the constrained solution space can make a difference.

% Instead of adaptively computing the weights, it adaptively selects them.

Synnev\aa{}g \emph{et al}. suggested using a window set comprised of rectangular, Kaiser (or Kaiser-Harris) and inverted Kaiser functions. In total he made LCA choose from 12 unique windows, 6 of which were phase-shifted versions of a fairly wide Kaiser window, and one were rectangular. Phase-shifting the windows has the effect of slighly adjusting the angle the array is steered towards. Synnev\aa{}g's composition of windows was motivated by the desire to use wide windows with maximum sidelobe suppression near the receiver, and narrow windows at greater depths for maximum sensitivity and depth penetration. While achieving good results, it was not made clear how many windows LCA needs, or how they ideally should be steered. Also, it is not apparent how LCA performs in sonar systems where the transmit beam is unfocused.

In this study we expand our previous work of applying LCA to active sonar imaging~\cite{Buskenes2011}. We use variations of the parametric Kaiser window, but also compare it to the trigonometric window function. We extend Synnev\aa{}g's work by steering all our windows and not just some, and demonstrate how LCA's performance is affected by the type and number of windows included. Our results are obtained using experimental data from the 32 element Kongsberg Maritime HISAS1030 sonar, which operates at 100\,kHz with an \minus{}3\,dB element opening angle of 23$^\circ$. Although this system was designed as synthetic aperture sonar (SAS) where an image is synthesized from several pings of data, we will only create images out of single pings (sectorscan imaging) in this study.

Our results show that when MVDR is modified to work in active sonar imaging, it computes weights with spatial responses that tend to be symmetric and steered within a fraction of the window's \minus{}3\,dB resolution. When letting LCA choose from Kaiser windows with similar responses, we obtain images with improved noise suppression and resolution. The resolution gain is made predictable by limiting the maximum steering to a fraction of the window's \minus{}3\,dB width. Overall, LCA produces images that are similar to MVDR, but with a resolution in between that of DAS and MVDR. LCA is also inherently robust, easy to implement, and fairly easy to understand.

This article is outlined as follows: In Section \ref{sec:beamforming} we offer a gentle introduction to beamforming. Then we move on to adaptive beamforming and describe the MVDR and LCA methods in \ref{sec:mvdr} and \ref{sec:lca}, respectively. We use LCA with the Kaiser window which we describe in \ref{sec:lca_kaiser_windows}, explain how we steer it in \ref{sec:lca_steering}, and by how much in \ref{sec:lca_steering_bounds}. Finally we add some remarks on using the trigonometric window instead of Kaiser in \ref{sec:lca_trigonometric}, and of of sampling considerations in \ref{sec:lca_oversampling}. Results and discussion are provided in Section \ref{sec:results_discussion}, where we study MVDR's windows to find suitable LCA ones in \ref{sec:results_lca_window_function}, assess the range of window types and steering in \ref{sec:results_window_parameters}, determine the number of windows needed in \ref{sec:results_database_size}, and add a note on computational complexity in \ref{sec:results_complexity}. Section \ref{sec:conclusion} provides a conclusion.

% This article is outlined as follows: In Section \ref{methods}, we introduce the concept of adaptive beamforming and provide details on the MVDR and LCA method. Then, in \ref{maptogpu}, we investigate the complexity issue, discuss means for reducing arithmetic complexity, and detail an implementation that makes efficient use of the GPU's parallel resources. The final design is assessed in Sections \ref{images_and_benchmarks} and \ref{discussion}, where we provide benchmarks, comparisons with similar CPU implementations, and measures of how efficiently our implementation makes use of the GPU's resources.

% - How much oversampling
% - How many parameters, how great a span.
% 
% - Experiementl data, Holmengraa/cross
% 
% - Note on efficiency.
% 
% 
% From our work with MVDR it 
% 
% 
% Traditionally beamformers were static, but as processors get more powerful so with the emergence of more processing power

% 
% \section{Methods}
% 
% At heart of the image formation process is a technique called beamforming, which creates a focus point at the pixel of interest. It does this by applying a suitable set of delays and weights to the array elements, such that signals emanating from the pixel location are emphasized constructively, while noise and interference from other directions sum destructively.
% 
% In active sonar array imaging, an acoustic wave is transmitted and the received echoes are usually recorded using an array of sensors. Each of these may be independently delayed and weighted such that signals emanating from directions of interest are added constructively, while noise and interference from other directions add destructively. Additionally, a window may be applied to the sensor channels to further adjust the array's spatial response. This process is known as beamforming.% The process of combining the signals from each sensor is known as beamforming.
% 
% 
% 
% 
% % Traditional beamformers such as Delay-and-Sum (DAS) apply predefined windows to all incoming data. However, due to the non-stationary nature of sonar data, the optimal window for any given time instant will generally differ from the next. This is where adaptive beamformers thrive, because they compute the optimal window coefficients for the data at each time instant. The choice of optimisation criteria is what mainly differentiates the various adaptive beamformers. The Minimum Variance (MV) beamformer, for instance, selects the set of weights that minimizes the beamformer output power for any given time instant~\cite{cap69}.
% 
% 
% \section{Background}
% 
% The MV beamformer computes the window coefficients by estimating and inverting a spatial covariance matrix. For improved estimation accuracy the data is averaged in space or time, or both. In addition, a certain amount of diagonal loading is often added to the covariance matrix before inversion. These actions, while leading to statistical robustness, tend to deteriorate the beamformer's performance. Also, the inversion of the covariance matrix has a computational complexity of O($M^3$), where $M$ is the number of elements. The computational burden makes the MV beamformer for large arrays less attractive, and sometimes even infeasible.
% 
% Based on the proposed method by Synnevåg~\cite{syn11}, we have implemented the LCA beamformer that keeps the minimum variance optimisation criterion but reduces the solution space to a discrete set of predefined windows. This reduces the computational complexity to O($MW$), where $W$ is the number of windows in the set. We use the Kaiser window function because it allows us to design a wide range of windows with different mainlobe widths and sidelobe suppression by adjusting the tradeoff parameter $\beta$~\cite{kai66}. In addition we apply steering to each of the windows and constrain the window design to ensure unit gain in the look direction.
% 
% The following will show that the proposed method performs similarly to or better than the MV method.

\section{Receive Beamforming}\label{sec:beamforming}

A sonar image is formed by estimating source locations and amplitudes. For this purpose a spatial bandpass filter is applied to the backscattered wavefield data. The filter is called an array processor or receive beamformer. The basic principle is to apply delay and weights to the sensor channels before summing them up, chosen such that signals from the location of interest are summed coherently, while other sources sum incoherently.

Assume that the wavefield has been sampled by an $M$ element uniform linear array, and that the signal signature has been removed by a matched filter. Let $x_m[\theta,n]$ be the delayed data from the $m$th channel, where the $\theta$ and $n$ are the azimuth angle and range sample of the focus point, respectively. Each angle $\theta$ will be processed independently, so to simplify notation we will assume the dependence on $\theta$ to be implicit from now on.

The output $z[n]$ of a beamformer is defined as the weighted sum of all the delayed data samples:
%
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]},\label{eq:beamformer_output}
\end{align}
%
where $w_m$ is the weight factor assigned to channel $m$. A weight set is commonly called a window, apodization or taper function. When the window is real its response is symmetric. Applying a real window that trails off towards the edges creates a response with lower sidelobe levels and wider mainlobe, which translates into improved noise suppression at the cost of reduced resolution, respectively~\cite{Harris1978}.

Conventional beamformers all have static and usually real weights. The reference method is the delay-and-sum (DAS) beamformer, also know as the backprojection algorithm. It delays each pixel into focus, then applies a suitable window, and finally sums the data. The virtue of DAS is its simplicity, robustness to parameter errors, linear processing of the image and the ease of which it can be implemented in parallel hardware.

Adaptive beamformers are dynamic and seek to adjust the array response to better fit the incoming wavefield. This may be achieved by allowing either the weights or delays to change, or both. Here the weights are usually complex, which allows asymmetric responses. One of the most extensively studied adaptive methods is the minimum variance distortion-less response (MVDR) beamformer.

\section{MVDR}\label{sec:mvdr}

MVDR seeks to minimize the output variance of the beamformer, under the constraint of a unity gain in some desired direction $\phi$~\cite{Capon1969}. With the assumption of zero mean data this can be formulated as: 
%
\begin{align}
\underset{\w[n]}{\argmin}\, E\{\big|z[n]\big|^2\} &= \underset{\w[n]}{\argmin}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w^H[n]\a_\phi &= 1.\label{eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. This is a convex optimization problem with the solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi}.\label{eq:mvdr_weights}
\end{gather}
The problem lies in estimating and inverting this spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009a}, and to improve robustness to parameter errors we add {\large$\epsilon$} percent diagonal loading~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

\begin{figure}[tbp]\centering%
\ifPeerReview%
\includegraphics[width=.5\linewidth]{gfx/buske1_online.eps}%
\else%
\includegraphics[width=\linewidth]{gfx/buske1_online.eps}%
\fi%
\caption{Each Kaiser window is steered in the interval $\phi=[0, \phi_{\mathrm{3\,dB}}(\beta)]$. The angle $\phi_{\mathrm{3\,dB}}(\beta)$ is the amount of steering needed for the steered response to have a \protect\minus{}3\,dB crossing that is exactly half that of the unsteered window. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.}\label{windows_steering}
\end{figure}

\begin{figure}[tbp]\centering%
\ifPeerReview%
\includegraphics[width=.5\linewidth]{gfx/buske2_online.png}%
\else%
\includegraphics[width=\linewidth]{gfx/buske2_online.png}%
\fi%
\caption{One of the objects used to test the beamformers is a 1\,m by 1\,m resolution test cross attached to an anchor with a diameter of approximately 13\,cm. Source image curtesy of Bundeswehr Technical Center for Ships and Naval Weapons, Maritime Technology and Research (WTD 71).}\label{cross}
\end{figure}

\setcounter{topnumber}{1}
\setcounter{dbltopnumber}{1}

\begin{figure*}[t]\centering%
\includegraphics[width=\linewidth]{gfx/buske3_online.pdf}%
\caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\newline
\emph{Left:}\hfill
\parbox[t]{.95\linewidth}{MVDR sectorscan image of the oiltanker Holmengraa, with the pixel groups are indicated with red boxes.}\protect\\\hspace{\textwidth}
\emph{Right:}\hfill
\parbox[t]{.95\linewidth}{MVDR window amplitude and phase responses computed from 40x40 pixel groups for the shadow, highlight and speckle region of the image. The responses are overlayed each other and the amount of overlap is colored using a logarithmic scale. The size of the pixel groups were chosen ad-hoc for the histograms to be near invariant to a shift in position within the same region. Since each pixel is pre-delayed into focus the unsteered responses all have their center at broadside. The dashed red lines at -13\,dB and -43\,dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3\,dB in highlight. The phase varies most in the highlight region where we see the highest contrast.} }\label{mvdr_selected_windows}
\end{figure*}

\begin{figure*}[t]%
\includegraphics[width=\textwidth]{gfx/buske4_online.pdf}%
\caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large and equal size window database, but where the upper bounds for $\beta$ and $\phi$ is varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
\newline
\emph{Upper bound $\beta$:}\hfill
\parbox[t]{.89\linewidth}{Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{mvdr_selected_windows} suggests that MVDR tend to choose windows with levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.}\newline
\emph{Upper bound $\phi$:}\hfill
\parbox[t]{.89\linewidth}{As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution. }
}\label{oversampling_mosaic_bounds}
\end{figure*}

% \setcounter{dbltopnumber}{2}

\begin{figure*}[t]\centering
\subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/buske5a_online.pdf}%%
\label{oversampling_mosaic_bounds_beta}}
\hfil
\subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/buske5b_online.pdf}%
\label{oversampling_mosaic_bounds_phi}}
\caption{Kaiser windows chosen for each pixel in the image. The underlying image is the one shown in \Fig{oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
\label{fig_sim}
\end{figure*}

\begin{figure*}[t]%\centering%
\includegraphics[width=\textwidth]{gfx/buske6_online.pdf}%
\caption{\emph{LCA images using different window database sizes}. The window boundaries are the same. The upper left image is identical to a rectangularly weighted DAS. Observe that sampling the $\beta$-range more finely increases the sidelobe suppression but leaves resolution unchanged. Adding more steering-variations improves both sidelobe suppression and resolution. However, using more than  $N_\beta=2$ window types and $N_\phi=3$ steering angles makes minimal difference.}\label{oversampling_mosaic}
\end{figure*}

\begin{figure*}[t]\centering
\subfloat[\emph{Image quality of approximate point scatterers found in the resolution test cross scene}. Two lateral image cuts are shown in the leftmost figures, along with the response of a rectangularly weighted DAS. Note how the FWHM of LCA sits in between that of DAS and MVDR. Also note how LCA is insensitive to the window type; its performance is near identical whether it uses Kasier or trigonometric windows.]{\includegraphics[width=\textwidth]{gfx/buske7a_online.pdf}%
\label{beamformer_comparison_cross}}
\hfil
\subfloat[\emph{Image quality of the full sector Holmengraa scene}. Compared to DAS the adaptive beamformers produce deeper shadows, sharper edges and a higher detail level. The adaptive methods produce nearly identical images.]{\includegraphics[width=\textwidth]{gfx/buske7b_online.pdf}%%
\label{beamformer_comparison_holmengraa}}
\caption{Comparing image quality of DAS, MVDR and LCA with Kaiser or trigonometric windows.}\label{image_quality}
\end{figure*}

% \begin{figure*}[t]%\centering%
% \includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_beta.pdf}%
% \caption{$\beta$-values of the windows LCA selects for each image pixel. The underlying image is the one shown in \Fig{oversampling_mosaic_bounds}, with the same parameters. The location of the anchor and cut line is marked in red. Note how LCA prefers a rectangular window ($\beta=0$) on the anchor.}\label{oversampling_mosaic_bounds_beta}
% \end{figure*} 
% 
% \begin{figure*}[t]%\centering%
% \includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_phi.pdf}%
% \caption{$\beta$-values of the windows LCA selects for each image pixel. The underlying image is the one shown in \Fig{oversampling_mosaic_bounds}, with the same parameters. The location of the anchor and cut line is marked in red. Note how LCA prefers a rectangular window ($\beta=0$) on the anchor.}\label{oversampling_mosaic_bounds_beta}
% \end{figure*}

% % \setcounter{dbltopnumber}{1

% \begin{figure*}[t]%\centering%
% \includegraphics[width=\textwidth]{gfx/beamformer_comparison_cross.pdf}%
% \caption{}\label{beamformer_comparison_cross}
% \end{figure*}
% 
% \begin{figure*}[t]%\centering%
% \includegraphics[width=\textwidth]{gfx/beamformer_comparison_holmengraa.pdf}%
% \caption{}\label{beamformer_comparison_holmengraa}
% \end{figure*}

\section{LCA}\label{sec:lca}

A much less complex alternative to MVDR is the low complexity adaptive (LCA) beamformer. It iterates through a set of $P$ windows and selects the window $p$ that best fulfills the minimum variance criterion:
%
\begin{align}
\underset{p}{\argmin}\, E\{\big|z_p[n]\big|^2\} = \underset{p}{\argmin}\, E\big\{\big|\w_{p}\H\x[n]\big|^2\big\}\nn
\qquad\text{subject to}\qquad \w\H\a_\phi = 1.\label{eq:lca_definition}
\end{align}\label{lca_criterion}
%
Note how closely LCA is related to the MVDR definition in \eq{eq:mvdr_definition}. The optimization criterion and constraint is the same, but LCA has a finite and discrete solution space for the weights. As will be demonstrated in section \ref{sec:results_lca_window_function}, LCA will perform similarly to MVDR because a robustified MVDR implementation seems to be constrained to window functions similar to the ones we let LCA choose from.

In practice, we estimate the beamformer output power by computing a sample power average $s_z^2$:
%
\begin{align*}
E\big\{\big|z[n]\big|^2\big\} \approx s_z^2 = \frac{1}{N_k} \sumb{n'=n-K}{n+K} \big| z[n'] \big|^2
\end{align*}
%
which computes the sample power average over $N_k=2K+1$ temporal samples. This is the same temporal averaging method we use for MVDR. In our case the bandwidth of our matched filtered signal is a just few samples long, hence $K=1$ will be used for both MVDR and LCA throughout this work.

If the pixels are correlated laterally, we can include a weighted combination of these as well to improve the variance estimation:
%
\begin{align}
s_z^2 = \frac{1}{N_x N_k} \sumb{\nmat{\scriptstyle x'=\\[-2mm]\scriptstyle x-X}}{x+X} \sumb{\nmat{\scriptstyle n'=\\[-2mm]\scriptstyle n-K}}{n+K} \omega[x',n']\big| z[x',n'] \big|^2
\end{align}
%
where $N_x = 2X+1$ is the number of azimuth lines to average over, and $\omega[x',n']$ is a normalized 2 dimensional weight function. Since we oversample slightly laterally when we delay each pixel (see Section \ref{sec:lca_oversampling}), we thought the lateral correlation to be sufficient to benefit from this. However, we observed no visual improvement from applying this technique, and decided not to use it.
% 
% \begin{itemize}
% \item \emph{Window function}
% \item \emph{Parameter boundaries}
% \item \emph{Number of windows needed}
% \item \emph{Oversampling need}
% \item \emph{Image quality}
% \end{itemize}

\subsection{Window function: Kaiser}\label{sec:lca_kaiser_windows}

As will be demonstrated in upcoming sections, the LCA beamformer works very well with windows generated from the Kaiser-Bessel function. We will express it in vector form as:
%
\begin{align}
\f_\beta = \bmat{
f_0(\beta) \\
\vdots\\
f_{M-1}(\beta)
}\label{eq:kaiser_window_vector}
\end{align}
%
where
%
\begin{align}
f_m(\beta) = \frac{I_0\left(\pi\beta\sqrt{1-\left(\frac{2m}{M-1}-1\right)^2}\right)}{I_0(\pi\beta)}\label{eq:kaiser_window_element}
\end{align}
%
and $I_0$ is the zeroth order modified Bessel function of the first kind:
%
\begin{align}
I_0(x) = \sumb{a=0}{\infty} \left[ \frac{\left(\frac{x}{2}\right)^a}{a!} \right]^2.\label{eq:modified_bessel_first_kind}
\end{align}
%
The Kaiser-Bessel window is near optimal in the sense of having its peak energy concentration around $\theta=0^\circ$, for a given space-bandwidth product related to the Kaiser parameter $\beta$ as:
%
\begin{align}
\beta = \frac{TB}{2},
\end{align}
%
where $T$ in our case is the spatial extent of the window and $B$ is its bandwidth. Adjusting $\beta$ changes the trade-off between mainlobe width and sidelobe level. When $\beta=0$ the window becomes rectangular, while at large values ($\beta>5$) the window converges to a Gaussian both in time and frequency. This class of windows is generally considered well suited for separating closely spaced sources with amplitudes of a high dynamic range~\cite{Harris1978}, they are easy to make, and they are optimal for any value of $\beta$.

\subsection{Steering}\label{sec:lca_steering}

Adding slightly steered versions of each window to the window database gives LCA greater flexibility in searching for an optimal window. A Kaiser window $\f_\beta$ steered to the angle $\phi$ can be expressed as 
%
\begin{align}
\w_{\beta,\phi} = \frac{\f_\beta\H\diag{(\a_\phi)}}{\f_\beta\H\a_\phi}\label{eq:steered_window}
\end{align}
%
where diag($\a_\phi$) is a diagonal matrix constructed from the steering vector $\a_\phi$:
%
\begin{align}
\a_\phi = \bmat{
1 \\
e^{-j\frac{2\pi d}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{2\pi (M-1)d}{\lambda}\sin(\phi)}
}.\label{eq:steering_vector}
\end{align}
%
Here $d$ is the element spacing, $\lambda$ is the wavelength and $\phi$ is the steering amount. The normalization factor $\f_\beta\H\a_\phi$ is the reciprocal of the window's coherent gain. It ensures unit gain in the direction of interest as required by (\ref{eq:lca_definition}). Since the signal-to-noise ratio is constant for a specific window (same $\beta$-value), this normalization also proportionally increases the incoherent noise gain.

The window database we will construct will contain $N_\beta$ Kaiser windows with unique $\beta$-values, each being steered in $N_\phi$ different directions. This gives us $N_w = N_\beta N_\phi$ unique windows.

\subsection{Steering bounds}\label{sec:lca_steering_bounds}

Each window's response is constrained to unit gain in the look direction. Hence, when it is steered the white noise gain must increase and the the signal-to-noise ratio decrease, as shown in \Fig{windows_steering}. To limit this we place an upper bound to the steering, chosen such that wide windows are allowed to be steered more than narrow ones. We call this upper bound for steering $\phi_{\mathrm{3\,dB}}(\beta)$, and define it as the steering angle needed for the steered response to have a \minus{}3\,dB crossing that is exactly half that of the unsteered window. The method we devised for finding it is described in Appendix \ref{appendix:steering_angles}.

Henceforth any steering will be specified relative to this upper bound. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.

\subsection{Trigonometric windows}\label{sec:lca_trigonometric}

While we focus on using LCA with Kaiser windows, we will for completeness compare the performance with that of using trigonometric windows. This is defined as:
%
\begin{align}
f_m(\alpha) = \alpha - (1-\alpha)\cos\Big(\frac{2\pi m}{M-1}\Big)\label{eq:trig_window_function}
\end{align}
%
where we apply the contraint $\alpha\in[0.5,1]$ to avoid windows with negative coefficients. Steering is applied as in (\ref{eq:steered_window}), and the $\alpha$-value that halves the \minus{}3\,dB distance is found using the method described in Appendix \ref{appendix:steering_angles}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%                      ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% IEEE APPROVED PACKAGES %
%%%%%%%%%%%%%%%%%%%%%%%%%%                      ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

@@

@@

\input{document}
